Blum axioms and nondeterministic computation of functions

### 2016-04-06 paper.tex
paper.tex: Improve the explanation of acceptable Gödel numberings.

One of the most important theoretical results concerning Turing machines
is the existence of undecidable problems.
Namely, the \emph{halting problem}
(the task of deciding whether a given Turing machine will halt on a given input)
cannot be solved by Turing machines.
To formalize (and prove) this fact,
we must first decide how to encode Turing machines in strings
(because all a Turing machine can manipulate are strings).

@@

In a way,
we are going the opposite direction:
instead of associating functions with strings,
we want to associate machines with functions.
We will use the concept of \emph{acceptable Gödel numbering}
(\cite[p.~41]{Rogers1987},~\cite[p.~324]{Blum1967})
to impose reasonableness criteria
to our definitions.

@@

Condition~\ref{universal-tm} is the universal Turing machine theorem.
By varying the first input of~$U$,
we can obtain any partial recursive function in a controlled manner.

And Condition~\ref{smn-theorem} is known as the $S_{mn}$ theorem,
or the iteration theorem.

>>>>

One of the most important theoretical results concerning Turing machines
is the existence of undecidable problems.
Namely, the \emph{halting problem}
(the task of deciding whether a given Turing machine will halt on a given input)
cannot be solved by Turing machines.
The formalization (and proof) of this fact
requires the definition of some sort of \emph{encoding};
since Turing machines can only reason about strings,
we need somehow to encode Turing machines into strings,
to be able to pose the halting problem as a language question.

@@

In a way,
we are going the opposite direction:
instead of associating functions with strings,
we want to associate machines with functions.

One important feature about the standard encodings of deterministic Turing machines
is the Universal Turing Machine Theorem:
the existence of a partial recursive function
$U: \{0, 1\}^* \times \{0, 1\}^* \to \{0, 1\}^*$
% Não sei se $U$ é um bom nome...
such that, if $w$ encodes the machine $M$,
then $U(m, x)$ is the result of running the machine $M$ on $x$.
The machines that compute $U$ are called \emph{universal Turing machines}.

The concept of \emph{acceptable Gödel numbering}
(\cite[p.~41]{Rogers1987},~\cite[p.~324]{Blum1967})
encompasses the existence of universal machines and a little more.
We will use it as our reasonableness criteria to our definitions.

@@

Condition~\ref{universal-tm} is the universal Turing machine theorem.

Condition~\ref{smn-theorem} is the ``little more'' we mentioned earlier.
It is known as the $S_{mn}$ theorem, or the iteration theorem.
Essentially,
given a partial recursive function of two variables,
we can obtain a partial recursive function of one variable
by fixing the first argument.
The function $\sigma$ provides a systematic way of doing this:
given a description $w$ of a partial recursive function of two variables
and the value $x$ to be fixed as the first variable,
the function returns a machine $\sigma(w, x)$
that computes the same function as $w$ but with the first argument fixed as $x$.
